Word Learning as Network Growth: A Cross-linguistic Analysis

### 2018-07-06 journal/ms.tex
journal version

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{Word Learning as Network Growth: A Cross-linguistic Analysis}

\usepackage[subtle]{savetrees}

\author{{\large \bf Abdellah Fourtassi} \\ \texttt{afourtas@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Yuan Bian} \\ \texttt{ybian.uiuc@gmail.com} \\ Department of Psychology \\ University of Illinois \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

\begin{document}

@@

\begin{abstract}
Children tend to produce words earlier when they are connected to a
variety of other words along both the phonological and semantic
dimensions. Though this connectivity effect has been extensively
documented, little is known about the underlying developmental
mechanism. One view suggests that learning is primarily driven by a
network growth model where highly connected words in the child's early
lexicon attract similar words. Another view suggests that learning is
driven by highly connected words in the external learning environment
instead of highly connected words in the early internal lexicon. The
present study tests both scenarios systematically in both the
phonological and semantic domains, and across 8 languages. We show that
external connectivity in the learning environment drives growth in both
the semantic and the phonological networks, and that this pattern is
consistent cross-linguistically. The findings suggest a word learning
mechanism where children harness their statistical learning abilities to
(indirectly) detect and learn highly connected words in the learning
environment.

\textbf{Keywords:}
semantic network, phonological network, network growth, mechanism of
word learning
\end{abstract}

@@

While most studies have focused on the static properties of the lexical
network, a few have investigated the underlying developmental process.
In particular, Steyvers \& Tenenbaum (2005) suggested that the observed
effects of connectivity are the consequence of how the lexical network
gets constructed in the child's mind. According to this explanation,
known as Preferential Attachment (PAT), highly connected words in the
child's lexicon tend to ``attract'' more words over time, in a
rich-get-richer scenario (Barabasi \& Albert, 1999). In other words,
what predicts word learning is the \emph{internal} connectivity in the
child's early lexicon. In contrast, Hills et al. (2009) suggested that
what biases the learning is not the connectivity in the child's internal
lexicon but, rather, \emph{external} connectivity in the learning
environment. They called this alternative explanation Preferential
Acquisition (PAC). Figure \ref{fig:growth} shows an illustration of both
growth scenarios with the same simplified network. These two proposals
represent two divergent ideas about the role of lexical networks in
acquisition. On the PAT proposal, network structure is a causal factor
in early word learning; in contrast, on the PAC approach, network
structure is not internally represented and, therefore, might be an
epiphenomenon of the statistics of the linguistic input.

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/growth-1}

@@

\caption{\label{fig:growth}Illustration of the growth scenarios. Filled circles (I1-I4) represent known words (internal), and empty circles (E1 and E2) represent words that have not been learned yet (external). Black lines represent links that are relevant in each growth scenario, and gray lines represent links that are irrelevant. For PAT, the utility of a candidate, external node is the average degree (i.e., number of links) of the internal nodes that it would attach to. Thus, according to PAT, the node E1 is more likely to enter the lexicon first. For PAC, the utility of a candidate node is its degree in the entire network. According to PAC, the node E2 is more likely to enter the lexicon first.}\label{fig:growth}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/all_data-1}

\caption{\label{fig:data_all}Age of acquisition in the global network as predicted by the degree in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits.}\label{fig:all_data}
\end{figure*}
\end{CodeChunk}

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/degree_distribution-1}

\caption{\label{fig:degree_distribution}Log-log plot of the cumulative degree distribution function for the global phonological and semantic networks across languages. A perfect power-law distribution should appear as a straight line in this graph.}\label{fig:degree_distribution}
\end{figure*}
\end{CodeChunk}

@@

\subsection{Semantic networks}\label{semantic-networks}

We constructed semantic networks following the procedure outlined in
Hills et al. (2009). We used as an index of semantic relatedness the
Florida Free Association Norms (Nelson, McEvoy, \& Schreiber, 1998).
This dataset was collected by giving adult participants a word (the
cue), and asking them to write the first word that comes to mind (the
target). For example, when given the word ``ball'', they might answer
with the word ``game''. A pair of nodes were connected by a directed
link from the cue to the target if there was a cue-target relationship
between these nodes in the association norms. The connectivity of a
given node was characterized by its \emph{indegree}: the number of links
for which the word was the target. To model growth from month to month,
we constructed a different network at each month, based on the words
that have been acquired by that month.

@@

@@

Connectivity in the global network is directly related to PAC as it
represents the explicit criterion PAC uses to determine what words
should be learned first (Figure \ref{fig:growth}). Therefore, a direct
consequence of a PAC-like growth scenario is a correlation between
connectivity in the global network and the age of
acquisition.\footnote{This correlation is also compatible with PAT, although the causality is reversed. Indeed, from the perspective of this growth scenario, higher connectivity in the global network is caused by earlier learning, not the other way around. Some words end up being highly connected in the global network precisely because they happen to be acquired earlier and, therefore, have a higher chance of accumulating more links over time.}
Figure \ref{fig:data_all} shows how the age of acquisition for each word
varies as a function of its degree (or indegree for the semantic
network). For ease of visual comparison, the predictor (i.e., the
degree) was centered and scaled across languages. The plots show,
overall, a negative correlation between the month of acquisition and the
degree, indicating that nouns with higher degrees are generally learned
earlier.

@@

We also analyzed the global network's degree distribution. The shape of
this distribution is particularly relevant to PAT as this growth
scenario is known to generate networks with a power-law degree
distribution (i.e., a distribution of the form
\(p(k) \propto \frac{1}{k^{\alpha}}\), Barabasi \& Albert, 1999). If the
network displays this property, this fact would suggest a PAT-like
generative process. Conversely, if the degree distribution does not
follow a power law, this fact would weaken the case for PAT. The log-log
plots are shown in Figure \ref{fig:degree_distribution}. We fit a power
law to each empirical degree distribution following the procedure
outlined in Clauset, Shalizi, \& Newman (2009) and using the related R
package (poweRlaw, Gillespie, 2015). In brief, the analysis consisted in
two steps. First, we derived the optimal cut-off, \(k_{min}\), above
which the distribution is more likely to follow a power
law,\footnote{In natural phenomena, it is often the case that the power law applies only for values above a certain minimum.}
and we estimate the corresponding scaling parameter \(\alpha\). Second
we calculated the goodness-to-fit, which resulted in a \(p\)-value
quantifying the plausibility of the model. Overall, we could not reject
the null hypothesis of a power-law distribution: the \(p\)-value was
generally above 0.1, except for the Italian phonological network where
we obtained \(p < 0.05\), suggesting that the power law can be ruled out
in this particular case.

@@

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/pred_ind_img-1}

\caption{\label{fig:pred_ind}Evaluation of network growth scenarios both individually (dotted), and when combined in the same growth model (solid). Each dot represents the mean of the posterior distribution of the corresponding growth parameter, with ranges representing 95\% credible intervals (computed using the highest density intervals). Positive values mean that learning proceeds according to the predictions of the growth scenario. Negative values mean that learning proceeds in opposition to the predictions of the growth scenario.}\label{fig:pred_ind_img}
\end{figure*}
\end{CodeChunk}

@@

We estimated the parameter \(\beta\) using a Bayesian approach. The
inference was performed using the probabilistic programming language
WebPPL (N. Goodman \& Stuhlmuller, 2014). We defined a uniform prior
over \(\beta\), and at each month, we computed the likelihood function
over words that could possibly enter the lexicon at that month, fit to
the words that have been learned at that month (using formula 1). Markov
Chain Monte Carlo sampling resulted in a posterior distribution over
\(\beta\), which we summarized in Figure \ref{fig:pred_ind}.

@@

@@

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/regressions_img-1}

\caption{\label{fig:regressions_img}Estimates of predictor coefficients by language, with ranges indicating 95\% confidence intervals. Positive values indicate a positive relationship (e.g. longer words tend to have a higher AoA), while negative values indicate a negative relationship (e.g. words with higher frequency tend to have a lower AoA).}\label{fig:regressions_img}
\end{figure*}
\end{CodeChunk}

@@

We used the frequency estimates from Braginsky et al. (2016) where
unigram counts were derived based on CHILDES corpora in each
language.\footnote{Note that these frequency counts are based on transcripts from independent sets of children and represent a general estimate of environmental frequency across children.}
For each word, counts included words that shared the same stem (e.g.,
``cats'' counts as ``cat''), or words that were synonymous (e.g.
``father'' counts as ``daddy''). For word length, we counted the number
of phonemes in our generated IPA transcription.

We conducted two analyses. We fit a linear regression for each language,
and we fit a linear mixed-effect model to all the data pooled across
languages, with language as a random effect. Figure
\ref{fig:regressions_img} shows the coefficient estimate for each
predictor in each language, and Figure \ref{fig:regressions_all_img}
shows the coefficient estimates for all languages combined (all
predictors were centered and scaled). The findings were as follows.
Overall, frequency is the largest and most consistent predictor of age
of acquisition, replicating results for nouns across a variety of
analyses (Braginsky et al., 2016; J. C. Goodman et al., 2008; B. C. Roy
et al., 2015). Word length predicts learning in some languages such as
Croatian and Norwegian, but not in others (including English). It
remains, however, a significant predictor in the global model. As for
the factors of interest, i.e., semantic and phonological connectivity,
we also found cross-linguistic differences. Phonological connectivity
contributes to learning in languages such as Croatian, English and
Russian, whereas semantic connectivity contributes to learning in
Turkish, Spanish and to some extent in Croatian, but not in
English.\footnote{Semantic connectivity does not explain variance in English data beyond that explained by phonological connectivity, frequency and length. This contrasts with the original finding in Hills et al. 2009. However, in this previous study, semantic connectivity was not tested in a model that included frequency, length and phonological connectivity as covariates. Another important difference is the number of words tested: Our study uses a larger set of nouns.}
Despite these cross-linguistic differences, both phonological and
semantic connectivity are significant predictors in the combined model.

@@

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/regressions_all_img-1}

\caption{\label{fig:regressions_all_img}Estimates of predictor coefficients in the combined mixed-effect model with language as a random effect. Ranges indicate 95\% confidence intervals. Lighter points indicate estimates of PAC predictors in a model that does not include frequency and length as covariates.}\label{fig:regressions_all_img}
\end{figure}
\end{CodeChunk}

@@

\fbox{\parbox[b][][c]{7.3cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/afourtassi/networks}}}
\vspace{1em}

@@

This work was supported by a post-doctoral grant from the Fyssen
Foundation, NSF \#1528526, and NSF \#1659585.

@@

\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent

@@

\hypertarget{ref-nelson1998}{}
Nelson, D. L., McEvoy, C. L., \& Schreiber, T. A. (1998). The University
of South Florida word association, rhyme, and word fragment norms.
Retrieved from \url{http://w3.usf.edu/FreeAssociation/}

>>>>

\documentclass[english,floatsintext,man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Word Learning as Network Growth: A Cross-linguistic Analysis},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

\ifxetex
  \usepackage{polyglossia}
  \setmainlanguage{}
\else
  \usepackage[english]{babel}
\fi

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

 % Line numbering
  \usepackage{lineno}
  \linenumbers

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{Word Learning as Network Growth: A Cross-linguistic Analysis}

  \shorttitle{Word Learning as Network Growth}

  \author{Abdellah Fourtassi\textsuperscript{1}, Yuan Bian\textsuperscript{2}, \& Michael C. Frank\textsuperscript{1}}

  % \def\affdep{{"", "", ""}}%
  % \def\affcity{{"", "", ""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{1} Department of Psychology, Stanford University\\
          \textsuperscript{2} Department of Psychology, University of Illinois  }

  \authornote{
    Abdellah Fourtassi

    Department of Psychology

    Stanford University

    50 Serra Mall

    Jordan Hall, Building 420

    Stanford, CA 94301

    Correspondence concerning this article should be addressed to Abdellah
    Fourtassi, Postal address. E-mail:
    \href{mailto:afourtas@stanford.edu}{\nolinkurl{afourtas@stanford.edu}}
  }

  \abstract{Children tend to produce words earlier when they are connected to a
variety of other words along both the phonological and semantic
dimensions. Though this connectivity effect has been extensively
documented, little is known about the underlying developmental
mechanism. One view suggests that learning is primarily driven by a
network growth model where highly connected words in the child's early
lexicon attract similar words. Another view suggests that learning is
driven by highly connected words in the external learning environment
instead of highly connected words in the early internal lexicon. The
present study tests both scenarios systematically in both the
phonological and semantic domains, and across 8 languages. We show that
external connectivity in the learning environment drives growth in both
the semantic and the phonological networks, and that this pattern is
consistent cross-linguistically. The findings suggest a word learning
mechanism where children harness their statistical learning abilities to
(indirectly) detect and learn highly connected words in the learning
environment.}
  \keywords{Language understanding; audio-visual processing; word learning; speech
perception; computational modeling. \\

  \usepackage[sortcites=false,sorting=none]{biblatex}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

@@

\setcounter{secnumdepth}{0}

@@

While most studies have focused on the static properties of the lexical
network, a few have investigated the underlying developmental process.
In particular, Steyvers and Tenenbaum (2005) suggested that the observed
effects of connectivity are the consequence of how the lexical network
gets constructed in the child's mind. According to this explanation,
known as Preferential Attachment (PAT), highly connected words in the
child's lexicon tend to \enquote{attract} more words over time, in a
rich-get-richer scenario (Barabasi \& Albert, 1999). In other words,
what predicts word learning is the \emph{internal} connectivity in the
child's early lexicon. In contrast, Hills et al. (2009) suggested that
what biases the learning is not the connectivity in the child's internal
lexicon but, rather, \emph{external} connectivity in the learning
environment. They called this alternative explanation Preferential
Acquisition (PAC). Figure~\ref{fig:growth} shows an illustration of both
growth scenarios with the same simplified network. These two proposals
represent two divergent ideas about the role of lexical networks in
acquisition. On the PAT proposal, network structure is a causal factor
in early word learning; in contrast, on the PAC approach, network
structure is not internally represented and, therefore, might be an
epiphenomenon of the statistics of the linguistic input.

\begin{figure}

{\centering \includegraphics[width=400px]{figs/growth2}

@@

\caption{Illustration of the growth scenarios. Filled circles (I1-I4) represent known words (internal), and empty circles (E1 and E2) represent words that have not been learned yet (external). Black lines represent links that are relevant in each growth scenario, and gray lines represent links that are irrelevant. For PAT, the utility of a candidate, external node is the average degree (i.e., number of links) of the internal nodes that it would attach to. Thus, according to PAT, the node E1 is more likely to enter the lexicon first. For PAC, the utility of a candidate node is its degree in the entire network. According to PAC, the node E2 is more likely to enter the lexicon first.}\label{fig:growth}
\end{figure}

@@

@@

\section{Semantic networks}\label{semantic-networks}

We constructed semantic networks following the procedure outlined in
Hills et al. (2009). We used as an index of semantic relatedness the
Florida Free Association Norms (Nelson, McEvoy, \& Schreiber, 1998).
This dataset was collected by giving adult participants a word (the
cue), and asking them to write the first word that comes to mind (the
target). For example, when given the word \enquote{ball}, they might
answer with the word \enquote{game}. A pair of nodes were connected by a
directed link from the cue to the target if there was a cue-target
relationship between these nodes in the association norms. The
connectivity of a given node was characterized by its \emph{indegree}:
the number of links for which the word was the target. To model growth
from month to month, we constructed a different network at each month,
based on the words that have been acquired by that month.

@@

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/corrAll-1} \caption{Age of production (A) and comprehension (B) in the global network as predicted by the degree in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits.}\label{fig:corrAll}
\end{figure}

@@

Connectivity in the global network is directly related to PAC as it
represents the explicit criterion PAC uses to determine what words
should be learned first (Figure \ref{fig:growth}). Therefore, a direct
consequence of a PAC-like growth scenario is a correlation between
connectivity in the global network and the age of
acquisition.\footnote{This correlation is also compatible with PAT, although the causality is reversed. Indeed, from the perspective of this growth scenario, higher connectivity in the global network is caused by earlier learning, not the other way around. Some words end up being highly connected in the global network precisely because they happen to be acquired earlier and, therefore, have a higher chance of accumulating more links over time.}
Figure \ref{fig:corrAll} shows how the age of production (A) and
comprehension (B) for each word varies as a function of its degree (or
indegree for the semantic network). For ease of visual comparison, the
predictor (i.e., the degree) was centered and scaled across languages.
The plots show, overall, a negative correlation between the month of
acquisition and the degree, indicating that nouns with higher degrees
are generally learned earlier.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{figs/plot_degree} \caption{Log-log plot of the cumulative degree distribution function for the global phonological and semantic networks across languages. The figure shows the results for both production and comprehension data. A perfect power-law distribution should appear as a straight line in this graph.}\label{fig:degreeDist}
\end{figure}

@@

We also analyzed the global network's degree distribution. The shape of
this distribution is particularly relevant to PAT as this growth
scenario is known to generate networks with a power-law degree
distribution (i.e., a distribution of the form
\(p(k) \propto \frac{1}{k^{\alpha}}\), Barabasi \& Albert, 1999). If the
network displays this property, this fact would suggest a PAT-like
generative process. Conversely, if the degree distribution does not
follow a power law, this fact would weaken the case for PAT. The log-log
plots are shown in Figure \ref{fig:degreeDist}. We fit a power law to
each empirical degree distribution following the procedure outlined in
Clauset, Shalizi, and Newman (2009) and using the related R package
(poweRlaw, Gillespie, 2015). In brief, the analysis consisted in two
steps. First, we derived the optimal cut-off, \(k_{min}\), above which
the distribution is more likely to follow a power
law,\footnote{In natural phenomena, it is often the case that the power law applies only for values above a certain minimum.}
and we estimate the corresponding scaling parameter \(\alpha\). Second
we calculated the goodness-to-fit, which resulted in a \(p\)-value
quantifying the plausibility of the model. The results are shown in
Appendix 1, table XX. Overall, we could not reject the null hypothesis
of a power-law distribution: the \(p\)-value was generally above 0.1.

@@

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/growthPred-1} \caption{Evaluation of network growth scenarios, for both A) production and B) comprehension. We plot values when predictors are evaluted individually (dotted), and when combined in the same growth model (solid). Each dot represents the mean of the posterior distribution of the corresponding growth parameter, with ranges representing 95\% credible intervals (computed using the highest density intervals). Positive values mean that learning proceeds according to the predictions of the growth scenario. Negative values mean that learning proceeds in opposition to the predictions of the growth scenario.}\label{fig:growthPred}
\end{figure}

@@

We estimated the parameter \(\beta\) using a Bayesian approach. The
inference was performed using the probabilistic programming language
WebPPL (N. Goodman \& Stuhlmuller, 2014). We defined a uniform prior
over \(\beta\), and at each month, we computed the likelihood function
over words that could possibly enter the lexicon at that month, fit to
the words that have been learned at that month (using formula 1). Markov
Chain Monte Carlo sampling resulted in a posterior distribution over
\(\beta\), which we summarized in Figure \ref{fig:growthPred}.

@@

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/staticPred-1} \caption{Estimates of predictor coefficients by language in production (A) and comprehension (B), with ranges indicating 95\% confidence intervals. Positive values indicate a positive relationship (e.g. longer words tend to have a higher AoA), while negative values indicate a negative relationship (e.g. words with higher frequency tend to have a lower AoA).}\label{fig:staticPred}
\end{figure}

@@

\begin{figure}[!h]
\includegraphics[width=\textwidth]{figs/plot_static_preds_all} \caption{Estimates of predictor coefficients in the combined mixed-effect model with language as a random effect for both production and comprehension. Ranges indicate 95\% confidence intervals. Lighter points indicate estimates of PAC predictors in a model that does not include frequency and length as covariates.}\label{fig:staticAll}
\end{figure}

@@

We used the frequency estimates from Braginsky et al. (2016) where
unigram counts were derived based on CHILDES corpora in each
language.\footnote{Note that these frequency counts are based on transcripts from independent sets of children and represent a general estimate of environmental frequency across children.}
For each word, counts included words that shared the same stem (e.g.,
\enquote{cats} counts as \enquote{cat}), or words that were synonymous
(e.g. \enquote{father} counts as \enquote{daddy}). For word length, we
counted the number of phonemes in our generated IPA transcription.

We conducted two analyses. We fit a linear regression for each language,
and we fit a linear mixed-effect model to all the data pooled across
languages, with language as a random effect. Figure \ref{fig:staticPred}
shows the coefficient estimate for each predictor in each language for
production and comprehension data. Figure \ref{fig:staticAll} shows the
coefficient estimates for all languages combined (all predictors were
centered and scaled). The findings were as follows. Overall, frequency
is the largest and most consistent predictor of age of acquisition,
replicating results for nouns across a variety of analyses (Braginsky et
al., 2016; J. C. Goodman et al., 2008; B. C. Roy et al., 2015). Word
length predicts learning in some languages such as Croatian and
Norwegian, but not in others (including English). It remains, however, a
significant predictor in the global model. As for the factors of
interest, i.e., semantic and phonological connectivity, we also found
cross-linguistic differences. Phonological connectivity contributes to
learning in languages such as Croatian, English and Russian, whereas
semantic connectivity contributes to learning in Turkish, Spanish and to
some extent in Croatian, but not in
English.\footnote{Semantic connectivity does not explain variance in English data beyond that explained by phonological connectivity, frequency and length. This contrasts with the original finding in Hills et al. 2009. However, in this previous study, semantic connectivity was not tested in a model that included frequency, length and phonological connectivity as covariates. Another important difference is the number of words tested: Our study uses a larger set of nouns.}
Despite these cross-linguistic differences, both phonological and
semantic connectivity are significant predictors in the combined model.

@@

@@

\fbox{\parbox[b][][c]{14cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/afourtassi/WordRec}}}
\vspace{1em}

@@

This work was supported by a post-doctoral grant from the Fyssen
Foundation.

\section{Disclosure statement}\label{disclosure-statement}

None of the authors have any financial interest or a conflict of
interest regarding this work and this submission.

\section{Appendix 1: Power law model}\label{appendix-1-power-law-model}

\begin{table}[H]
\centering
\begin{tabular}{rrrrll}
  \hline
 & Kmin & alpha & pValue & dimension & language \\ 
  \hline
1 & 4.00 & 2.55 & 0.88 & Sem & Croatian \\ 
  2 & 4.00 & 2.18 & 0.12 & Phono & Croatian \\ 
  3 & 4.00 & 2.38 & 0.00 & Sem & Danish \\ 
  4 & 11.00 & 4.55 & 0.86 & Phono & Danish \\ 
  5 & 5.00 & 2.66 & 0.13 & Sem & English \\ 
  6 & 20.00 & 9.14 & 0.51 & Phono & English \\ 
  7 & 8.00 & 2.81 & 0.13 & Sem & French \\ 
  8 & 20.00 & 3.75 & 0.11 & Phono & French \\ 
  9 & 4.00 & 2.93 & 0.61 & Sem & Italian \\ 
  10 & 9.00 & 9.45 & 0.78 & Phono & Italian \\ 
  11 & 5.00 & 2.88 & 0.20 & Sem & Norwegian \\ 
  12 & 15.00 & 6.28 & 0.74 & Phono & Norwegian \\ 
  13 & 24.00 & 5.61 & 0.72 & Sem & Russian \\ 
  14 & 8.00 & 4.20 & 0.54 & Phono & Russian \\ 
  15 & 4.00 & 2.98 & 0.46 & Sem & Spanish \\ 
  16 & 13.00 & 8.75 & 0.74 & Phono & Spanish \\ 
  17 & 4.00 & 2.49 & 0.17 & Sem & Swedish \\ 
  18 & 11.00 & 4.68 & 0.10 & Phono & Swedish \\ 
  19 & 4.00 & 2.87 & 0.93 & Sem & Turkish \\ 
  20 & 8.00 & 3.26 & 0.38 & Phono & Turkish \\ 
   \hline
\end{tabular}
\caption{Results of fitting a power law model to the degree distribution in each model for production data. Kmin is the optimal degree cut-off, alpha is the scaling parameter, and pValue is the probability that quantifies the plausibility of the power law hypothesis. If pValue is close to 1, the power law model cannot be rejected as a plausible fit for the data. If, instead, pValue is small (e.g., p < 0.05) then the null hypothesis of a power law model can be rejected.} 
\end{table}\begin{table}[H]
\centering
\begin{tabular}{rrrrll}
  \hline
 & Kmin & alpha & pValue & dimension & language \\ 
  \hline
1 & 5.00 & 2.67 & 0.90 & Sem & Croatian \\ 
  2 & 2.00 & 2.06 & 0.02 & Phono & Croatian \\ 
  3 & 4.00 & 2.39 & 0.01 & Sem & Danish \\ 
  4 & 5.00 & 2.98 & 0.14 & Phono & Danish \\ 
  5 & 4.00 & 2.64 & 0.77 & Sem & English \\ 
  6 & 13.00 & 5.16 & 0.23 & Phono & English \\ 
  7 & 4.00 & 2.63 & 0.33 & Sem & French \\ 
  8 & 18.00 & 5.58 & 0.34 & Phono & French \\ 
  9 & 4.00 & 2.88 & 0.69 & Sem & Italian \\ 
  10 & 8.00 & 10.27 & 0.91 & Phono & Italian \\ 
  11 & 5.00 & 2.87 & 0.43 & Sem & Norwegian \\ 
  12 & 13.00 & 7.65 & 0.44 & Phono & Norwegian \\ 
  13 & 8.00 & 3.91 & 0.95 & Sem & Russian \\ 
  14 & 5.00 & 3.97 & 0.85 & Phono & Russian \\ 
  15 & 5.00 & 3.11 & 0.55 & Sem & Spanish \\ 
  16 & 5.00 & 3.01 & 0.09 & Phono & Spanish \\ 
  17 & 5.00 & 2.81 & 0.71 & Sem & Swedish \\ 
  18 & 9.00 & 6.75 & 0.10 & Phono & Swedish \\ 
  19 & 4.00 & 3.13 & 0.89 & Sem & Turkish \\ 
  20 & 9.00 & 5.73 & 0.96 & Phono & Turkish \\ 
   \hline
\end{tabular}
\caption{Results of fitting a power law model to the degree distribution in each model for comprehension data. Kmin is the optimal degree cut-off, alpha is the scaling parameter, and pValue is the probability that quantifies the plausibility of the power law hypothesis. If pValue is close to 1, the power law model cannot be rejected as a plausible fit for the data. If, instead, pValue is small (e.g., p < 0.05) then the null hypothesis of a power law model can be rejected.} 
\end{table}

@@

\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

@@

\hypertarget{ref-nelson1998}{}
Nelson, D. L., McEvoy, C. L., \& Schreiber, T. A. (1998). \emph{The
University of South Florida word association, rhyme, and word fragment
norms}. Retrieved from \url{http://w3.usf.edu/FreeAssociation/}
